Reasoning about a Capability Machine with Local Capabilities

### 2017-06-14 popl-2018/paper.tex
Changed all permissions to use small caps - no bar
typo

% Permissions
\newcommand{\plainperm}[1]{\mathrm{#1}}
\newcommand{\noperm}{\plainperm{o}}
\newcommand{\readonly}{\plainperm{ro}}
\newcommand{\readwrite}{\plainperm{rw}}
\newcommand{\exec}{\plainperm{rx}}
\newcommand{\entry}{\bar{\plainperm{e}}}
\newcommand{\rwx}{\plainperm{rwx}}
% PWL permissions
\newcommand{\readwritel}{\plainperm{rwl}}
\newcommand{\rwl}{\readwritel}
\newcommand{\rwlx}{\plainperm{rwlx}}

% Global/local
\newcommand{\local}{\plainperm{local}}
\newcommand{\glob}{\plainperm{global}}

@@

%\dominique{``end up being defined recursively'' does not sound very insightful}
A region defines what its part of memory can legally look like. However,
validity of the memory's contents may itself depend on the state of other parts
of memory. To accommodate this, worlds need to be defined recursively: Worlds
describe memory, which can contain capabilities, whose safety is
world-dependent. Recursive worlds are a common requirement in Kripke models and
the following lemma uses the method of
\citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for
constructing them. The formulation of the lemma is technical, so we recommend
that non-expert readers ignore the technicalities and accept that there exists a
set of worlds $\Wor$ and two relations $\futurestr$ and $\futurewk$ satisfying
the (recursive) equations in the theorem (where the $\blater$ operator can be
safely ignored).
% Short technical explanation
% The recursiveness of the worlds gives rise to a recurisve domain
% equation. Theorem~\ref{thm:world-existence} gives the solution to this
% recursive domain equation.
% \dominique{it is not explained what cofes are?}
% \lau{I added what cofe means but no formal definition (I think this goes under the category technicalities)}
\begin{theorem}\label{thm:world-existence}
  There exists a \cofe{} (complete ordered family of equivalences) $\Wor$ and preorders $\futurestr$ and
  $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are
  preordered \cofes{}, and there exists an isomorphism $\xi$ such that
  \begin{align*}
      \xi \; : \; & \Wor \cong \blater (\nats \finparfun \Regions)\\
      \Regions  = & \; \{\revoked\} \uplus \\
               & \; \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments}))\uplus \\
               & \; \{\perma\} \times \States \times \Rels \times (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
    \end{align*}
  and for $W, W' \in \Wor$
  \begin{align*}
    W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
    W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
  \end{align*}
\end{theorem}
In the above theorem, $\States \times \Rels$ corresponds to the
aforementioned state transition system where $\Rel$ is a set of pairs
of relations corresponding to the public and private transition, and
$\States$ is a sufficiently large set of states containing at least the
states we use. The last part of the temporary and permanent regions is
a state interpretation function that determines what memory segments
the region permits in each state of its states. The different
monotonicity requirements in the two interpretation functions reflects
how permanent regions rely only on permanent protocols whereas
temporary regions can rely on both temporary and permanent protocols.
$\UPred{\HeapSegments}$ is a step-indexed, downwards closed predicate
on memory segments.

>>>>

% Permissions
\newcommand{\plainperm}[1]{\textsc{#1}}
\newcommand{\noperm}{\plainperm{o}}
\newcommand{\readonly}{\plainperm{ro}}
\newcommand{\readwrite}{\plainperm{rw}}
\newcommand{\exec}{\plainperm{rx}}
\newcommand{\entry}{\plainperm{e}}
\newcommand{\rwx}{\plainperm{rwx}}
% PWL permissions
\newcommand{\readwritel}{\plainperm{rwl}}
\newcommand{\rwl}{\readwritel}
\newcommand{\rwlx}{\plainperm{rwlx}}

% Global/local
\newcommand{\plainlocality}[1]{\mathrm{#1}}
\newcommand{\local}{\plainlocality{local}}
\newcommand{\glob}{\plainlocality{global}}

@@

%\dominique{``end up being defined recursively'' does not sound very insightful}
A region defines what its part of memory can legally look like. However,
validity of the memory's contents may itself depend on the state of other parts
of memory. To accommodate this, worlds need to be defined recursively: Worlds
describe memory, which can contain capabilities, whose safety is
world-dependent. Recursive worlds are a common requirement in Kripke models and
the following lemma uses the method of
\citet{Birkedal:2011:SKM:1926385.1926401,Birkedal:tutorial-notes} for
constructing them. The formulation of the lemma is technical, so we recommend
that non-expert readers ignore the technicalities and accept that there exists a
set of worlds $\Wor$ and two relations $\futurestr$ and $\futurewk$ satisfying
the (recursive) equations in the theorem (where the $\blater$ operator can be
safely ignored).
% Short technical explanation
% The recursiveness of the worlds gives rise to a recurisve domain
% equation. Theorem~\ref{thm:world-existence} gives the solution to this
% recursive domain equation.
% \dominique{it is not explained what cofes are?}
% \lau{I added what cofe means but no formal definition (I think this goes under the category technicalities)}
\begin{theorem}\label{thm:world-existence}
  There exists a \cofe{} (complete ordered family of equivalences) $\Wor$ and preorders $\futurestr$ and
  $\futurewk$ such that $(\Wor,\futurestr)$ and $(\Wor,\futurewk)$ are
  preordered \cofes{}, and there exists an isomorphism $\xi$ such that
  \begin{align*}
      \xi \; : \; & \Wor \cong \blater (\nats \finparfun \Regions)\\
      \Regions  = & \; \{\revoked\} \uplus \\
               & \; \{\temp\} \times \States \times \Rels \times (\States \fun (\Wor \monwknefun \UPred{\HeapSegments}))\uplus \\
               & \; \{\perma\} \times \States \times \Rels \times (\States \fun (\Wor \monstrnefun \UPred{\HeapSegments}))
    \end{align*}
  and for $W, W' \in \Wor$
  \begin{align*}
    W' \futurestr W & \Leftrightarrow \xi(W') \futurestr \xi(W)   \\
    W' \futurewk W & \Leftrightarrow \xi(W') \futurewk \xi(W)
  \end{align*}
\end{theorem}
In the above theorem, $\States \times \Rels$ corresponds to the
aforementioned state transition system where $\Rels$ is a set of pairs
of relations corresponding to the public and private transition, and
$\States$ is a sufficiently large set of states containing at least the
states we use. The last part of the temporary and permanent regions is
a state interpretation function that determines what memory segments
the region permits in each state of its states. The different
monotonicity requirements in the two interpretation functions reflects
how permanent regions rely only on permanent protocols whereas
temporary regions can rely on both temporary and permanent protocols.
$\UPred{\HeapSegments}$ is a step-indexed, downwards closed predicate
on memory segments.
